Just Take the Bus: Analysis of Toronto Bike Theft from 2014-2023

### 2024-11-13 Paper/Article.Rmd
repositioned graphs

knitr::opts_chunk$set(echo=FALSE, include=TRUE, eval=TRUE)
required_packages <- c(
  "lubridate",
  "sf",
  "tidyverse",
  "ggplot2",
  "maps",
  "dplyr",
  "mapdata",
  "ggspatial",
  "opendatatoronto",
  "here",
  "stringr",
  "RColorBrewer",
  "Scales",
  "pak",
  "patchwork"
) 
for (p in required_packages) { # install if missing
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(lubridate)
library(sf)
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(mapdata)
library(ggspatial)
library(opendatatoronto)
library(here)
library(stringr)
library(RColorBrewer)
library(scales)
library(pak)
library(patchwork)
```

@@

```{r}
thefts_by_neighborhood <- bike_thefts %>%
  count(NEIGHBOURHOOD_Clean) %>%
  rename(thefts_count = n)

@@

 ggplot(toronto_thefts_map) +
  geom_sf(aes(fill = thefts_count), color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "Number of Thefts") +
  labs(title = "Number of Bike Thefts per Neighborhood in Toronto") +
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5))

bike_thefts %>%
  count(PREMISES_TYPE) %>%
   ggplot(aes(x = "", y = n, fill = PREMISES_TYPE)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(title = "Bike Theft by Location Type",
       fill = "Location Type") +
  theme_void()

# Display the plots side by side
#(plot_1 | plot_spacer() | plot_2) + 
 # plot_layout(widths = c(1, 0.3, 1))

@@

bike_thefts %>%
  count(OCC_MONTH) %>%
  ggplot(aes(x = OCC_MONTH, y = n, fill = OCC_MONTH)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Thefts by Month of the Year",
       x = "Month", y = "Number of Thefts", fill = "Months") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

@@

bike_thefts %>%
  count(OCC_DOW) %>%
  ggplot(aes(x = OCC_DOW, y = n, fill = OCC_DOW)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Theft Patterns by Week Day Slight Uptick on Fridays",
       x = "Day of the Week", 
       y = "Number of Thefts", 
       fill = "Day of Week",
       caption = "Bike thefts in Toronto show a consistent pattern across weekdays, with a small increase on Fridays.") +  
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")

@@

ggplot(recovery_status, aes(x = "", y = n, fill = STATUS)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5),
            color = "white") +
  labs(title = "Bike Recovery Status: Recovered vs Stolen",
       fill = "Status") +
  theme_void() +
  scale_fill_brewer(palette = "Dark2")

>>>>

knitr::opts_chunk$set(echo=FALSE, include=TRUE, eval=TRUE)
required_packages <- c(
  "lubridate",
  "sf",
  "tidyverse",
  "ggplot2",
  "maps",
  "dplyr",
  "mapdata",
  "ggspatial",
  "opendatatoronto",
  "here",
  "stringr",
  "RColorBrewer",
  "Scales",
  "pak",
  "patchwork",
  "gridExtra"
) 
for (p in required_packages) { # install if missing
  if (!require(p, character.only = TRUE)) {
    install.packages(p, character.only = TRUE)
  }
}

library(lubridate)
library(sf)
library(tidyverse)
library(ggplot2)
library(maps)
library(dplyr)
library(mapdata)
library(ggspatial)
library(opendatatoronto)
library(here)
library(stringr)
library(RColorBrewer)
library(scales)
library(pak)
library(patchwork)
library(gridExtra)
```

@@

```{r}
plot_1<- thefts_by_neighborhood <- bike_thefts %>%
  count(NEIGHBOURHOOD_Clean) %>%
  rename(thefts_count = n)

@@

 plot_1<-ggplot(toronto_thefts_map) +
  geom_sf(aes(fill = thefts_count), color = "grey", size = 0.1) +
  scale_fill_viridis_c(option = "plasma", name = "No. of Thefts") +
  labs(caption = "Thefts per Neighborhood in Toronto") +
  theme_void() +
  theme(plot.caption = element_text(hjust = 0.5))

plot_2<- bike_thefts %>%
  count(PREMISES_TYPE) %>%
   ggplot(aes(x = "", y = n, fill = PREMISES_TYPE)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  labs(caption = "Thefts by Location Type",
       fill = "Location Type") +
  theme_void()

grid.arrange(plot_1, plot_2, ncol=2, widths = c(2, 2))

@@

bike_thefts %>%
  count(OCC_MONTH) %>%
  ggplot(aes(x = OCC_MONTH, y = n, fill = OCC_MONTH)) +
  geom_bar(stat = "identity") +
  labs(caption = "Bike Thefts by Month of the Year",
       x = "Month", y = "Number of Thefts") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = "none") +
  theme(plot.caption = element_text(hjust = 0.5))
```

@@

bike_thefts %>%
  count(OCC_DOW) %>%
  ggplot(aes(x = OCC_DOW, y = n, fill = OCC_DOW)) +
  geom_bar(stat = "identity") +
  labs(title = "Bike Theft Patterns by Week Day Slight Uptick on Fridays",
       x = "Day of the Week", 
       y = "Number of Thefts", 
       fill = "Day of Week",
       caption = "Bike thefts in Toronto show a consistent pattern across weekdays, with a small increase on Fridays.") +  
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")+
  guides(fill = "none")

@@

ggplot(recovery_status, aes(x = "", y = n, fill = STATUS)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")), 
            position = position_stack(vjust = 0.5),
            color = "white") +
  labs(title = "Bike Recovery Status: Recovered vs Stolen",
       fill = "Status") +
  theme_void() +
  scale_fill_brewer(palette = "Dark2") +
  theme(title = element_text(hjust = 0.5, size = 12))
